## Exercise 16, p. 195: logistic regression only.

```{r}
library(ISLR)
Boston = read.csv("/Volumes/work/MTH522/data/Boston.csv")
head(Boston)
```

### Adding the crim01 variable into the dataset to identify if the crime rate is above or below the median(crim).

```{r}
Boston1 <- Boston
Boston1$crim01 <- NA
crim_median = median(Boston1$crim)
print(crim_median)

for(i in 1:dim(Boston1)[1]){
  if (Boston1$crim[i] > crim_median){
    Boston1$crim01[i] = 1
  }else{
    Boston1$crim01[i] = 0
  }
}
head(Boston1)
```

```{r}
pairs(Boston1[,1:15], main="Scatterplot matrix including all of the variables")
```

```{r}
autocorr = cor(Boston1$crim01,Boston1)
autocorr
```

**Observations:**

1\. From the above table we can observe that, `indus`,`nox`,`age`,`rad`,`tax` are the variables that are statistically significant to `crim01`

```{r}
require(caTools)
set.seed(123)
Boston_split = sample.split(Boston1$crim01, SplitRatio = 0.80)
Boston_train = subset(Boston1, Boston_split==TRUE) 
Boston_test = subset(Boston1, Boston_split==FALSE)
```

**Observations:**

1\. logistic regression for the variables which are statistically significant to `crim01`. i.e.,`indus`,`nox`,`age`,`rad` and `tax`

```{r}
Boston1_glm = glm(crim01 ~ indus+nox+age+rad+tax, data=Boston_train, family=binomial)
Boston1_prob = predict(Boston1_glm,Boston_test,type="response")
Boston1_pred = rep(0,length(Boston_test$crim01))
Boston1_pred[Boston1_prob >0.5] = 1

table(Boston1_pred,Boston_test$crim01)
```

```{r}
mean(Boston1_pred != Boston_test$crim01)
```

**Observations:**

1\. The logistic model has a 11.76% error rate. It means it has a correct prediction of 88.24%
